Doubly-Robust Inference in R using \\texttt{drtmle}

### 2022-06-24 paper/using_drtmle_obstudies.Rmd
fix tex template

---
output:
  latex_fragment:
    includes:
      in_header: preamble.tex
bibliography: refs.bib
---

@@

This has led to increased interest in adopting adaptive estimation techniques
from the machine learning literature to control for high-dimensional confounders
when estimating treatment effects. However, facilitating proper inference (i.e.,
confidence intervals and hypothesis tests) about estimates of effects based on
adaptive estimators is challenging. In particular, the use of standard causal
inference techniques, including both G-computation (GCOMP) estimators
[@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
estimators [@horvitzthompson:1952:jasa, @robins:2000:epi] is challenging when
utilizing adaptive estimators. In particular, with rare exception, if
cross-validation is used to select tuning parameters for the nuisance
parameters, estimates of the population-level effect will generally be too
biased to achieve standard $n^{1/2}$-rate asymptotics. Instead, *undersmoothing*
must be used to select tuning parameters to control the bias
[@undersmoothed_hal]. However, this can prove quite challenging in practice and
few guidelines are available.

@@

As an intermediate step, the AIPTW and TML estimators require estimation of two
key nuisance parameters: the probability of treatment as a function of
confounders (the propensity score, hereafter referred to as the *PS*) and the
average outcome as a function of confounders and treatment (the outcome
regression, *OR*). In addition to their desirable inferential properties, AIPTW
and TML estimators also enjoy a *double robustness* property. That is, the
estimators are consistent for the population-level effect of interest if either
of the PS or OR is consistently estimated. This property gives the AIPTW and TML
estimators a natural appeal: inconsistency in the estimation of one nuisance
parameter may be mitigated by the consistent estimation of the other. However,
recent work has shown that the double robustness property does not necessarily
extend to the inferential properties of these estimators when adaptive
estimators of the PS and/or OR are used
[@vdl:2014:ijb, @benkeser:2017:biometrika]. Instead, valid inference requires
consistent estimation of *both* the PS *and* the OR. @vdl:2014:ijb proposed new
TML estimators that have an asymptotic normal sampling distribution with
estimable variance under consistent estimation of *either* the PS *or* the OR,
paving the way for inference that is doubly-robust. @benkeser:2017:biometrika
proposed modifications of these estimators with similar robustness properties
and illustrated their practical performance.

@@

```{r, echo = FALSE, message = FALSE}
library(SuperLearner)
```
Super learning is a loss-based ensemble machine learning technique
[@vdl:2007:statapp] that generalizes stacking and stacked regression
\citep{wolpert:1992:nnet,breiman:1996:ml}. The user specifies many potential
candidate regression estimators, referred to as a *library* of candidate
estimators, and the super learner algorithm estimates the convex combination of
regression estimators that minimizes the $V$-fold cross-validated risk based on
a user-selected loss function [@dudoit2005asymptotics].  Oracle inequalities
show that the super learner estimate performs essentially as well as the unknown
best convex combination of the candidate estimators [@vdvaart:2006:statdec].
Thus, the methodology may be seen as an asymptotically optimal way of performing
estimator selection. Practically, the super learner provides the best chance
for obtaining a consistent estimator for both the OR and PS, and thereby an
efficient estimator of the marginal means of interest.

@@

## References

>>>>

---
output:
  bookdown::pdf_document2:
    template: obs_study_skeleton.tex
bibliography: refs.bib
abs_text: |
  Inverse probability of treatment weighted estimators and doubly-robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss-based estimators) are widely used in causal inference
  to estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence
  intervals and p-values may be computed based on standard asymptotic results,
  such as the central limit theorem, the delta method, and the nonparametric
  bootstrap. However, in high-dimensional settings maximum likelihood estimation
  often breaks down and standard procedures no longer yield correct inference.
  Instead, we may rely on adaptive estimators of nuisance parameters to
  construct flexible regression estimators. However, use of adaptive estimators
  poses a challenge for performing statistical inference about an estimated
  treatment effect. While doubly-robust estimators facilitate inference when
  *all* relevant regression functions are consistently estimated, the same
  cannot be said when at least one nuisance estimator is inconsistent.
  \texttt{drtmle} implements doubly-robust confidence intervals and hypothesis
  tests for targeted minimum loss-based estimates of the average treatment
  effect, in addition to several other recently-proposed estimators of the
  average treatment effect.
abs_keywords: |
  causal inference, machine learning, semiparametric estimation, treatment
  effects, multiple robustness, targeted minimum loss estimation
---

@@

This has led to increased interest in adopting adaptive estimation techniques
from the machine learning literature to control for high-dimensional confounders
when estimating treatment effects. However, facilitating proper inference (i.e.,
confidence intervals and hypothesis tests) about estimates of effects based on
adaptive estimators is challenging. In particular, the use of standard causal
inference techniques, including both G-computation (GCOMP) estimators
[@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
estimators [@horvitzthompson:1952:jasa; @robins:2000:epi] is challenging when
utilizing adaptive estimators. In particular, with rare exception, if
cross-validation is used to select tuning parameters for the nuisance
parameters, estimates of the population-level effect will generally be too
biased to achieve standard $n^{1/2}$-rate asymptotics. Instead, *undersmoothing*
must be used to select tuning parameters to control the bias
[@undersmoothed_hal]. However, this can prove quite challenging in practice and
few guidelines are available.

@@

As an intermediate step, the AIPTW and TML estimators require estimation of two
key nuisance parameters: the probability of treatment as a function of
confounders (the propensity score, hereafter referred to as the *PS*) and the
average outcome as a function of confounders and treatment (the outcome
regression, *OR*). In addition to their desirable inferential properties, AIPTW
and TML estimators also enjoy a *double robustness* property. That is, the
estimators are consistent for the population-level effect of interest if either
of the PS or OR is consistently estimated. This property gives the AIPTW and TML
estimators a natural appeal: inconsistency in the estimation of one nuisance
parameter may be mitigated by the consistent estimation of the other. However,
recent work has shown that the double robustness property does not necessarily
extend to the inferential properties of these estimators when adaptive
estimators of the PS and/or OR are used
[@vdl:2014:ijb; @benkeser:2017:biometrika]. Instead, valid inference requires
consistent estimation of *both* the PS *and* the OR. @vdl:2014:ijb proposed new
TML estimators that have an asymptotic normal sampling distribution with
estimable variance under consistent estimation of *either* the PS *or* the OR,
paving the way for inference that is doubly-robust. @benkeser:2017:biometrika
proposed modifications of these estimators with similar robustness properties
and illustrated their practical performance.

@@

```{r, echo = FALSE, message = FALSE}
library(SuperLearner)
```
Super learning is a loss-based ensemble machine learning technique
[@vdl:2007:statapp] that generalizes stacking and stacked regression
[@wolpert:1992:nnet; @breiman:1996:ml]. The user specifies many potential
candidate regression estimators, referred to as a *library* of candidate
estimators, and the super learner algorithm estimates the convex combination of
regression estimators that minimizes the $V$-fold cross-validated risk based on
a user-selected loss function [@dudoit2005asymptotics]. Oracle inequalities
show that the super learner estimate performs essentially as well as the unknown
best convex combination of the candidate estimators [@vdvaart:2006:statdec].
Thus, the methodology may be seen as an asymptotically optimal way of performing
estimator selection. Practically, the super learner provides the best chance
for obtaining a consistent estimator for both the OR and PS, and thereby an
efficient estimator of the marginal means of interest.

@@

# References
